\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-22 pldi/draft.tex
More eval

At a high level, we are interested in understanding whether \sys can indeed scale shell scripts out automatically and correctly.
To answer this question, we use several scripts collected out from the wild, along with a few micro-benchmarks for targeted comparisons.

@@

\heading{Setup}
Measurements were taken on a machine with 512GB of memory and 64 physical $\times$ 2.1GHz Intel Xeon E5-2683 cores running Debian 4.9.144-3.1, GNU Core-utils 8.30-3, GNU Bash 5.0.3(1)-, Python 3.7.3, and OCaml 4.05.0.
No special configuration was made in hardware or software.
Except as otherwise noted,
  (i) all pipelines are set to (initially) read from and (finally) write to the file-system,
  (ii) \ttt{curl} fetches data from a different physical host on the same network connected by 1Gbps links.
% A more detailed description of the setup and experiments, including reproducible scripts, is included in the accompanying (anonymized) online repository.

@@

\heading{Discussion}
\kk{Below follows an important point about why scalability is more
  than great in the beginning and then gets worse.}
An interesting point is that \sys seems to achieve almost linear
speedup with small parallelism configurations. The reason why this
happens is not so straightforward though. \sys manages that because it
exploits task-based parallelism by having different processes to do
the \todo{map} and the \todo{reduce} phase. Due to task-based
parallelism in the generated dataflow graphs, \sys essentially ``uses
more parallelism'' than the given configuration. This can be seen in
all pipelines that are not completely stateless. For example, in the
\ttt{minimal\_sort} one-liner, \sys configured to have 8 parallelism
spawns 37 nodes (8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m}
nodes, and 14 \ttt{eager} nodes. This indicates that \sys achieves
optimal performance (without unnecessary process spawns and context
switches) in pipelines that contain pure commands by configuring it
with a smaller parallelism than the capabilities of the underlying
system -- in our case 16-32 for a 64 physical core system. This is
also the case if the input pipeline is very long and so there is
already task-based parallelism even in the sequential execution
(several of the Unix50 pipelines fall in this category).

\kk{Hypothesis on a scalability upper bound. Dish gives a great
  speedup if one step of the pipeline is what takes the most time. If
  the pipeline is long with many computation heavy tasks, task based
  parallelism gets you part of the way there, so we can't expect Dish
  to give performance benefits equal to the amount of parallelism that
  we give it.}

\TODO{Briefly discuss the performance drop as we increase parallelism
  in some one-liners (sort-sort, bi-grams).The reason has to do with
  context switching. More precisely, sort-sort 32 has 314 nodes (and
  54m sys time) and sort-sort 64 has 634 nodes (and 90m! sys time). On
  the other hand opt-bigrams 64 has 255 nodes and the sys time stays
  the same so I think that it might have to do with the fact that
  constant costs rise a lot (since opt-bigrams 32 takes just 1
  minute).}

\kk{It is visible that the automatic split is slightly worse than the
  manual one in most cases (because it doesn't utilize task-based
  parallelism).}

We attribute the small drop above $96\times$ in \ttt{spell} \ttt{sort} \ttt{wf}
and \ttt{bi-grams} to a combination of hyperthreading (128 virtual vs. 64
physical CPUs) \todo{the fixed width on} \ttt{sort}.
% I say "we attribute" because it's a way to say something without being certain
% FIXME: We need to take linear benchmarks on deathstar and livestar to see
% where it starts dropping
% not explain scope the results appropriately
% rinard

@@

\heading{Programs} In a recent event celebrating \unix's 50-year
legacy, Bell Labs created a set of \todo{36} challenges~\cite{unix50}
solvable by means of composing \unix pipelines.  By scanning GitHub,
we were able to find a set of shell programs that solve
\todo{all-but-three} problems~\cite{}.  We consider this a good set of
benchmarks because (i) the problems were designed to to highlight the
\unix philosophy~\cite{} (rather than having a single command that
solves the problem) and make extensive use of \unix built-ins with a
plethora of flags, and (ii) the solutions were written by non-experts
(contrary to the \S\ref{ours}) and were executed by \sys
unmodified---stressing patterns that we, \sys's authors, did not
necessarily anticipate \kk{I would change that to ``stressing a
  variety of patterns'' since it doesn't really matter if we
  anticipated them (especially since we can address any issues and
  then re-run them).} .  In cases where an obvious fix improves the
pipeline's performance, we report on both \sys's original speedup and
the one achieved by applying the fix.
% TODO: talk about correct wrt developer intentions vs. correct wrt problem definition

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}

\heading{Results} \sys's parallelism set to 16. The input given to the
pipelines was the test input from unix50 multiplied to reach
10GBs. The speedup of each pipeline separately is shown in
\Cref{fig:unix50-individual}. The average\todo{/median} speedup is
$5.49$/\todo{$6.07\times$}, the \todo{geometric mean is $1.08\times$},
and the \todo{weighted average (with the absolute times as weights) is
  $5.75\times$}.

@@

%% Breakdown of not ideal speedups in 8, 16 parallelism

%% Head: 2, 19, 31
%% Sort-bottleneck: 1, 3, 15, 16, 20, 33
%% Non-parallelizable completely: 13 (awk-solved), 18 (awk), 24, 25, 26, 29, 30
%% Cheap (non cpu intensive -- constant costs grow): 0, 4, 5, 6, 7, 8, 12, 14
%% Long (task-based parallelism in sequential and context switches): 7, 8, 9, 10, 11, 12, 14, 21, 27, 28

\sys achieves speedup in all pipelines except for 2, 19, and 31. These
pipelines include a \ttt{head} and only process one line of the
input. This can be seen by their sequential execution time which is in
the order of 10 milliseconds. The slowdown is due to constant costs of
setting up the pipes and spawning processes. In all three of these
cases the execution time with \sys is in the order of 1 second. For
pipelines 13, 24, 25, 26, 29, and 30, \sys performs the same as the
sequential implementation. This is because these pipelines contain
commands that can not be parallelized in general, such as \ttt{awk},
\ttt{sed} with option \ttt{d}, and \ttt{tr} with option \ttt{-d}. An
illustrative example is pipeline 13, which is shown here:

\begin{lstlisting}[language=sh, float=h, numbers=none]
  cat $IN | awk "{print \$2, \$0}" |
      sort -nr | cut -d ' ' -f 2
\end{lstlisting}

\kk{Rephrase that to sound more meaningful and surprising.}  Notice
that \ttt{awk} is just used to permute two of the columns of the input
to enable sorting. One could write this pipeline in a different way,
by directly sorting on the second column:

\begin{lstlisting}[language=sh, float=h, numbers=none]
  cat $IN6 | sort -nr -k 2 | cut -d ' ' -f 1
\end{lstlisting}

By executing this version of the pipeline with \sys, we achieve
\todo{X} speedup. This indicates that just a little awaresess of
command parallelizability can go a long way, and that the effort that
a \sys user needs to spend is often negligible to achieve
benefits.

@@

>>>>

This section reports on whether \sys can indeed scale shell scripts out automatically and correctly, using several scripts collected out from the wild along with a few micro-benchmarks for targeted comparisons.

@@

\heading{Setup}
\sys was run on 512GB of memory and 64 physical $\times$ 2.1GHz Intel Xeon E5-2683 cores, Debian 4.9.144-3.1, GNU Core-utils 8.30-3, GNU Bash 5.0.3(1)-, Python 3.7.3, and OCaml 4.05.0---without any special configuration in hardware or software.
Except as otherwise noted,
  (i) all pipelines are set to (initially) read from and (finally) write to the file-system,
  (ii) \ttt{curl} fetches data from a different physical host on the same network connected by 1Gbps links.
% A more detailed description of the setup and experiments, including reproducible scripts, is included in the accompanying (anonymized) online repository.

@@

\heading{Discussion}
\sys seems to achieve near-linear speedup on low-parallelism configurations.
This is due to a high degree of task-based parallelism from distinct \emph{map} and \emph{fold} processes, which results in a longer tree of task-parallel stages.
This is seen in all that include components in \pur.
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.

Thus, asking for parallelism beyond that point results in worse speedup, due to other overheads that start dominating as no further speedup is available.
Sort-sort with $32\times$ has 314 nodes (and 54m sys time) and with $64\times$ has 634 nodes (and 90m! sys time).
 On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
  the same so I think that it might have to do with the fact that
  constant costs rise a lot (since opt-bigrams 32 takes just 1
  minute).

%> \kk{It is visible that the automatic split is slightly worse than the
%>   manual one in most cases (because it doesn't utilize task-based
%>   parallelism).}

%> This is also the case if the input pipeline is very long and so there is already task-based parallelism even in the sequential execution (several of the Unix50 pipelines fall in this category).
%> \kk{Hypothesis on a scalability upper bound. Dish gives a great
%>   speedup if one step of the pipeline is what takes the most time. If
%>   the pipeline is long with many computation heavy tasks, task based
%>   parallelism gets you part of the way there, so we can't expect Dish
%>   to give performance benefits equal to the amount of parallelism that
%>   we give it.}

@@

\heading{Programs}
In a recent celebration of \unix's 50-year legacy, Bell Labs created 37 challenges~\cite{unix50} solvable by means of composing \unix pipelines.
We found unofficial solutions to all-but-three problems on GitHub~\cite{unix50sol}, expressed as pipelines with 2--12 stages (avg.: 5.58).
We consider this a good set of benchmarks because
  (i) the problems were designed to highlight the \unix modular philosophy~\cite{mcilroy1978unix} and make extensive use of built-ins with a variety of flags, and
  (ii) the solutions were written by non-experts (contrary to \S\ref{ours}).
% ---stressing a patterns that we, \sys's authors, did not
% necessarily anticipate \kk{I would change that to ``stressing a
%   variety of patterns'' since it doesn't really matter if we
%   anticipated them (especially since we can address any issues and
%   then re-run them).} . 
\sys executes each pipeline as-is; 
  if an obvious fix improves performance, we report the speedup of both versions. %  original and modified. %  pipeline.
% TODO: talk about correct wrt developer intentions vs. correct wrt problem definition

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}
\heading{Results}
\sys's parallelism was set to $16\times$, as many of these pipelines are relatively long (\S\ref{ours}, Discussion).
Their input was generated by multiplying the original Unix50 input several times to reach 10GB.
Fig~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines.
The average speedup is $5.49$ ($6.07\times$), and the weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\heading{Discussion}
Most pipelines see significant acceleration---except 13, 24, 25, 26, 29, and 30 that see no speedup and 2, 19, and 31 that see a slowdown.
Ones in the first group contain stages that cannot be parallelized in general---\eg \ttt{awk}, \ttt{sed} \ttt{-d}, and \ttt{tr} \ttt{-d}.
However, by applying trivial changes
that change some of these commands with their equivalents in \sta---\eg, replacing a general \ttt{awk} with more specialized \ttt{cut} that achieves the same effect---\sys 
can identify and exploit parallelism:
  for, say, \todo{13} this leads  to \todo{XX}$\times$ speedup.
This is a common pattern we witnessed elsewhere too~\sx{macro1}.

Ones in the second group contain a \ttt{head}, thus always processing only one line and executing within $10ms$ regardless of the full input.
\sys's slowdown is due to constant setup costs---\eg pipes, processes---but remains within $1s$.

For pipelines  \sys performs the same as the sequential implementation.

% illustrative example is pipeline 13, which is shown here:
% 
% \begin{lstlisting}[language=sh, float=h, numbers=none]
%   cat $IN | awk "{print \$2, \$0}" |
%       sort -nr | cut -d ' ' -f 2
% \end{lstlisting}
% 
% \kk{Rephrase that to sound more meaningful and surprising.}  Notice
% that \ttt{awk} is just used to permute two of the columns of the input
% to enable sorting. One could write this pipeline in a different way,
% by directly sorting on the second column:
% 
% \begin{lstlisting}[language=sh, float=h, numbers=none]
%   cat $IN6 | sort -nr -k 2 | cut -d ' ' -f 1
% \end{lstlisting}

@@

%% Breakdown of not ideal speedups in 8, 16 parallelism
%% Head: 2, 19, 31
%% Sort-bottleneck: 1, 3, 15, 16, 20, 33
%% Non-parallelizable completely: 13 (awk-solved), 18 (awk), 24, 25, 26, 29, 30
%% Cheap (non cpu intensive -- constant costs grow): 0, 4, 5, 6, 7, 8, 12, 14
%% Long (task-based parallelism in sequential and context switches): 7, 8, 9, 10, 11, 12, 14, 21, 27, 28
